Design and Implementation of Probabilistic Programming Language Anglican

### 2016-08-08 paper/paper.tex
minor

The sole purpose of the algorithm identifier of \texttt{infer} is to invoke
an appropriate overloading or implementation of the function --- conventionally, 
the identifier is a keyword related to the
algorithm name (\texttt{:lmh} for Lightweight
Metropolis-Hasting, \texttt{:pcascade} for Particle Cascade
etc.). The second parameter is a query as defined by
the \texttt{defquery} form or its anonymous version \texttt{query}. For instance,
the following Clojure code invokes \texttt{infer} on an Anglican query
defined anonymously via the \texttt{query} form:
%If needed, the query
%can be defined anonymously, right in the argument list of a call
%to \texttt{infer}:
\begin{lstlisting}[style=default]
(let [x 1]
  (infer :pgibbs (query (predict x)) nil))
\end{lstlisting}
A query is executed by calling the initial continuation of the
query, which accepts a value and a state. The state is supplied
by the inference algorithm, while the value is provided as a
parameter of \texttt{infer}. A query does not have to have any
parameters, in which case the value can be simply \texttt{nil}.
When a query is defined with a binding for the initial value,
the value becomes available inside the query. A query may
accept multiple parameters using Clojure's structured binding.
For instance, it may take multiple parameters as components 
of an input vector. In this case, the initial value is given
as a structured value, such as a vector, and the components
of this value gets pattern matched to the corresponding
parameters of the query via the destructuring mechanism of Clojure:
%The value is destructured via pattern match in case of a structured binding. 
For example,
\begin{lstlisting}[style=default]
(defquery my-query [mean sd]
  (predict (sample (normal mean sd))))

@@

The simplest inference algorithm is \textit{importance
sampling}:
\begin{lstlisting}[style=default]
(derive ::algorithm
        :anglican.inference/algorithm)
(defmethod infer :importance [_ prog value & {}]
  (letfn
    [(sample-seq []
       (lazy-seq
         (cons
           (:state (exec ::algorithm prog value
                         initial-state))
           (sample-seq))))]
    (sample-seq)))
\end{lstlisting}
\texttt{anglican.importance/{\linebreak[0]}infer} just calls
\texttt{anglican.inference/{\linebreak[0]}exec} and relies on default
implementations of checkpoint handlers. A different inference
algorithm would provide its own implementations of
\texttt{checkpoint} for \texttt{sample}, \texttt{observe}, or
both, as well as invoke \texttt{exec} from an elaborated
conditional control flow. LMH (\texttt{anglican.lmh}) and SMC
(\texttt{anglican.smc}) are examples of inference algorithms
where either \texttt{observe} (SMC) or \texttt{sample} (LMH)
handler is overridden. In addition, SMC runs multiple particles
(program instances) simultaneously, while LMH runs re-runs
programs from an intermediate continuation rather than from the
beginning.

>>>>

The sole purpose of the algorithm identifier of \texttt{infer} is to invoke
an appropriate overloading or implementation of the function --- conventionally, 
the identifier is a keyword related to the
algorithm name (\texttt{:lmh} for Lightweight
Metropolis-Hasting, \texttt{:pcascade} for Particle Cascade
etc.). The second parameter is a query as defined by
the \texttt{defquery} form or its anonymous version \texttt{query}. For instance,
the following Clojure code invokes \texttt{infer} on an Anglican query
defined anonymously via the \texttt{query} form:
%If needed, the query
%can be defined anonymously, right in the argument list of a call
%to \texttt{infer}:
\begin{lstlisting}[style=default]
(let [x 1]
  (infer :pgibbs (query (predict x)) nil))
\end{lstlisting}
A query is executed by calling the initial continuation of the
query, which accepts a value and a state. The state is supplied
by the inference algorithm, while the value is provided as a
parameter of \texttt{infer}. A query does not have to have any
parameters, in which case the value can be simply \texttt{nil}.
When a query is defined with a binding for the initial value,
the value becomes available inside the query. A query may
accept multiple parameters using Clojure's structured binding.
For instance, it may take multiple parameters as components 
of an input vector. In this case, the initial value is given
as a structured value, such as a vector, and the components
of this value gets pattern matched to the corresponding
parameters of the query via the destructuring mechanism of Clojure.
%The value is destructured via pattern match in case of a structured binding. 
For example,
\begin{lstlisting}[style=default]
(defquery my-query [mean sd]
  (predict (sample (normal mean sd))))

@@

The simplest inference algorithm is \textit{importance
sampling}:
\begin{lstlisting}[style=default]
(derive ::algorithm
        :anglican.inference/algorithm)
(defmethod infer :importance [_ prog value & {}]
  (letfn
    [(sample-seq []
       (lazy-seq
         (cons
           (:state (exec ::algorithm prog value
                         initial-state))
           (sample-seq))))]
    (sample-seq)))
\end{lstlisting}
\texttt{anglican.importance/{\linebreak[0]}infer} just calls
\texttt{anglican.inference/{\linebreak[0]}exec} and relies on default
implementations of checkpoint handlers. A different inference
algorithm would provide its own implementations of
\texttt{checkpoint} for \texttt{sample}, \texttt{observe}, or
both, as well as invoke \texttt{exec} from an elaborated
conditional control flow. LMH (\texttt{anglican.lmh}) and SMC
(\texttt{anglican.smc}) are examples of inference algorithms
where either \texttt{observe} (SMC) or \texttt{sample} (LMH)
handler is overridden. In addition, SMC runs multiple particles
(program instances) simultaneously, while LMH re-runs
programs from an intermediate continuation rather than from the
beginning.

\hy{This subsection is also not quite self-contained. Someone
who is not familiar with Clojure or Anglican codebase is likely
to get lost. On the other hand, it is a good overview or guide
for those who want to study the implemenation of Anglican. One
way to address this presentation issue while keeping the benefits
of the current text is to include some extended examples. For instance,
in the beginning of the paper, we might include several
Anglican examples. Then, we show that how these examples get
cps-transformed. Next, we present the implementation of inference
algorithms in Anglican using semi pseudo code; this is not quite
an implementation in Clojure but gives a good sketch on how
various functions are implemented in each inference algorithm.
Finally, we show how these algorithms work on cps-tranformed 
examples.}

\hy{I rewrote various parts so as to avoid the word multimethod.
My strategy was to use the phrases, such as (ad-hoc) polymophic
function and overloading. All of my rephrasing should be
double-checked.}
